### 2018-06-26 paper/drafts/litrev.tex
Changed the file structure

\documentclass{article}
\begin{document}

HI THIS IS THE START OF THE LIT REV
\section{Introduction}

Control systems on multirotor aerial vehicles have had many different iterations, each of which have added a significant level of control to the system. The control mechanism can be as simple as a motor-rotor control, all the way to controlling the torque, or thrust or so on. The control of a multirotor propellor has to account for multiple different factors, from the ambient wind speed, the rotoation, the pitch of the propellor, all the way to the high level control system of the ESC.

Our paper will compare our system with those of existing systems. 
Our system, a control system with a thrust sensor which provides a feed back loop for the system to check if the appropriate thrust is sought after.

\end{document}